---
title: Chapter 10 Homework Answers
hw_number: 5
pubdate: '2023-03-02'
date: '2023-03-01'
pdf_url: /files/homework_solutions/hw_05/homework_05_ch_10_answers.pdf
slug: homework_05_ch_10_answers
output:
  blogdown::html_page:
    md_extensions: +tex_math_single_backslash+compact_definition_lists
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 3.0
---
```{r setup, include=FALSE}
knitr::knit_hooks$set(inline = function(x) { knitr:::format_sci(x, 'md')})
knitr::opts_chunk$set(echo = FALSE, include = TRUE, cache = FALSE)

if (knitr::is_latex_output()) {
  message("Setting PDF plot hook.")
  new_plot_hook <- function(x, options)
    paste("\n", knitr::hook_plot_tex(x, options), "\n")

  knitr::knit_hooks$set(plot = new_plot_hook)
} else {
  message("Not using PDF output.")
}


options(dplyr.summarise.inform = FALSE)

# This section loads necessary R libraries and sources scripts that define
# useful functions format_md.
#
data_dir = "_data"
script_dir = "_scripts"

library(pacman)
p_load(tidyverse, jsonlite, httr, magrittr, scales, xml2, knitr)

theme_set(theme_bw(base_size = 10))

# Create a data directory if one does not exist.
if (!dir.exists(data_dir)) dir.create(data_dir)
if (!dir.exists(file.path(data_dir, "ch_10")))
    dir.create(file.path(data_dir, "ch_10"))


source(file.path(script_dir, "utils.R"), chdir = T)
source(file.path(script_dir, "format_md.R"), chdir = T)
source(file.path(script_dir, "geocarb.R"), chdir = T)
```
## Exercise 10.1: Long-term fate of fossil-fuel CO~2~

> Use the online geologic carbon cycle model at
> <https://climatemodels.jgilligan.org/geocarb/>.
> Use the default setup of the model, and notice that the CO~2~ weathering
> rates, etc., for the transient state are the same as for the spin-up state.
> So, if there were no CO~2~ spike at all, there would be no change in
> anything at year 0. (Go ahead, make sure I am not lying about this.)
> Release some CO~2~ in a transition spike, 1000 Gton or more or less,
> and see how long it takes for the CO~2~ to decrease to a plateau.
> There are two CO~2~ plots in the output, one covering 100,000 years and one
> covering 2.5 million years. How long does it take for CO~2~ to level out
> after the spike, according to both plots?

### Answer to 10.1

Even after 1 million years (the longest we can plot with the new version of
the Geocarb model), CO~2~ hasn't quite returned to its original value.

At the end of spinup, CO~2~ was 272.6 parts per million, and 1 million years
after the spike, it's 273.1 ppm.

## Exercise 10.4:

> Using the SLUGULATOR model at <https://climatemodels.jgilligan.org/slugulator/> or
> <http://climatemodels.uchicago.edu/slugulator/>,
> compare the impact of methane and CO~2~ on timescales of 1, 10, 25, 50, 100,
> 500, and 1000 years.
>
> The SLUGULATOR model simulates releasing a large amount of CO~2~ and
> methane and then calculates what happens over time. At time zero, the
> concentration of each gas is the natural level from shortly before the
> industrial revolution: 280 ppm of CO~2~ and 1.6 ppm of methane. At one year,
> a large amount of each gas is released instantly (after this, there are no
> further emissions of either gas).
>
> a) Run the model using the default model input parameters, setting the left
>    graph to "Concentrations" and the right graph to "Surface T Anomaly".
>    Set the time scale on the bottom to show 10 years.
>
>    Move your mouse over the graphs to measure the exact values (a "tooltip"
>    will pop up next to the cursor telling the time in years and the value of
>    the line on the graph: concentration of CO~2~ or methane, or the warming
>    due to that gas). Be aware that on the concentration graph, the
>    concentration of CO~2~ is plotted against the left axis while methane is
>    plotted against the right axis.
>
>    What happens to the concentrations of methane and carbon dioxide? Why does
>    the concentration of CO~2~ continue to rise after the first year, when
>    methane is falling?

### Answer to 10.4(a)

After the spike, methane immediately begins to drop, and by about 35 or 40 years
later, it's pretty much back to where it was before the spike.

After the spike, CO~2~ continues to rise for 35--40 years, and then gradually
falls over the next million years.

When methane falls during the first 50 years or so, it is removed from the 
atmosphere by chemical reactions that oxidize it and convert it to CO~2~
and water vapor. CO~2~ continues to rise after the initial spike because
the initial spike of methane is being converted to CO~2~.

> b) What is the warming due to methane at 1 year and at 10 years? What is the
>    warming due to CO~2~ at 1 and 10 years? (Use the mouse to examine the
>    time-dependent temperature plotted on the graphs, not the "Time-integrated
>    temperature" in the table at the top of the page). Calculate the ratio of
>    methane-warming to CO~2~-warming at each time.


### Answer to 10.4(b)

After 1 year, the warming due to methane is 0.029 degrees C. After 10 years,
it's 0.09 degrees.

After 1 year, the warming due to CO~2~ is 0.001 degree, and after 10 years,
it's 0.006 degrees.

At 1 year, the ratio of methane warming to CO~2~ warming is 
$0.029 / 0.001 = `r round(0.029 / 0.001)`$. 
At 10 years, the ratio is  
$0.09 / 0.006 = `r round(0.09 / 0.006)`$. 

> c) Change the time scale to show 25, 50, 100, 500, and 1000 years. Make a
>    table showing the warming due to each gas and the ratio of methane-warming
>    to CO~2~-warming at 1, 10, 25, 50, 100, 500, and 1000 years.

### Answer to 10.4(c)

```{r answer-4c}
#| echo=FALSE, warning=FALSE, message=FALSE
df <- tibble(
  year = c(1, 10, 25, 50, 100, 500, 1000),
  ch4 = c(0.029, 0.09,  0.03, 0.002, 0.001, 0,     0),
  co2 = c(0.001, 0.006, 0.01, 0.01,  0.01,  0.009, 0.007)
) %>%
  mutate(ratio = ch4 / co2)

df %>% 
  select(year, ch4, co2, ratio) %>%
  knitr::kable(col.names = c(year = "Year", ch4 = "Methane warming", 
                             co2 = "CO2 warming", ratio = "Ratio CH4/CO2"))
```

> d) Plot the ratio of methane-warming to CO~2~-warming over time.

### Answer to 10.4(d)

```{r answer-4d}
#| echo=FALSE, warning=FALSE, message=FALSE
ggplot(df, aes(x = year, y = ratio)) + 
  geom_point(size = 3, color = "darkblue") + 
  geom_line(size = 1, color = "darkblue") +
  labs(x = "Year", y = expression(paste("Ratio ", CH[4] / CO[2])))
```

> e) Why does the ratio change with time?

Because methane is a much more powerful greenhouse gas than CO~2~, but 
it has a much shorter lifetime in the atmosphere, so over time, the methane
goes away while the CO~2~ remains for thousands of years.

### Answer to 10.4(e)

Because methane is a much more powerful greenhouse gas than CO~2~, but 
it has a much shorter lifetime in the atmosphere, so over time, the methane
goes away while the CO~2~ remains for thousands of years.

> f) Generating electricity by burning natural gas (methane) releases less than
>    half as much CO~2~ than generating the same amount of electricity by
>    burning coal.
>
>    "Fracking" has dramatically lowered the cost of natural gas, and this has
>    made coal-fired electrical generation uneconomical, so about half of the
>    coal generation plants in the United States have shut down in the last 10
>    years.
>
>    Many people think replacing coal generation with natural gas generation
>    will reduce CO~2~ emissions significantly,
>    but gas wells that use fracking also leak a lot of methane directly into
>    the atmosphere.
>
>    Comment on the significance of this homework exercise to the debate
>    over whether fracking is helping or hurting the problem of global warming.

### Answer to 10.4(f)

In terms of long-term warming, the move from burning coal to burning methane
is a big improvement because less than half as much CO~2~ will be 
released into the atmosphere. 

However, in the short term (the next 50 years or so), the leaking methane from
the gas wells will produce a lot of warming. It's possible that enough 
methane could leak out of the wells to create more warming in the short-term
than the coal would have done.

However, this methane will disappear from the atmosphere in the next several
decades, whereas the CO~2~ from burning coal would remain in the atmosphere
for many thousands of years.
