---
title: Chapter 7 Homework Answers
hw_number: 4
pubdate: '2023-02-28'
date: '2023-02-22'
pdf_url: /files/homework_solutions/hw_04/homework_04_answers.pdf
slug: homework_04_answers
output:
  blogdown::html_page:
    md_extensions: +tex_math_single_backslash+compact_definition_lists
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 3.0
---
```{r setup, include=FALSE}
knitr::knit_hooks$set(inline = function(x) { knitr:::format_sci(x, 'md')})
knitr::opts_chunk$set(echo = FALSE, include = TRUE, cache = FALSE)

if (knitr::is_latex_output()) {
  message("Setting PDF plot hook.")
  new_plot_hook <- function(x, options)
    paste("\n", knitr::hook_plot_tex(x, options), "\n")

  knitr::knit_hooks$set(plot = new_plot_hook)
} else {
  message("Not using PDF output.")
}


options(dplyr.summarise.inform = FALSE)

# This section loads necessary R libraries and sources scripts that define
# useful functions format_md.
#
data_dir = "_data"
script_dir = "_scripts"

library(pacman)
p_load(tidyverse, jsonlite, httr, magrittr, scales, xml2, knitr)

theme_set(theme_bw(base_size = 10))

# Create a data directory if one does not exist.
if (!dir.exists(data_dir)) dir.create(data_dir)
if (!dir.exists(file.path(data_dir, "ch_07")))
    dir.create(file.path(data_dir, "ch_07"))


source(file.path(script_dir, "utils.R"), chdir = T)
source(file.path(script_dir, "format_md.R"), chdir = T)
source(file.path(script_dir, "modtran.R"), chdir = T)
source(file.path(script_dir, "rrtm.R"), chdir = T)
```
## **Exercise #1: Cloud Effects**

> Use the MODTRAN model with the "Tropical Atmosphere" and vary the
> clouds.
> Note that this model does not consider incoming shortwave light, and
> therefore it ignores the albedo effect of clouds. It only calculates the
> effect of the clouds on outgoing longwave light for a given surface
> temperature. One way to think about this is that it looks at the effect the
> clouds have at night.
>
> a) Run the model three times: First with clear skies, then with
>    "Altostratus: Cloud Base 2.4 km, Top 3.0 km,"
>    and finally with "Stratus: Cloud Base .33 km, Top 1.0 km."
>    Describe the change in \(I_{\text{out}}\) for each type of clouds:
>
>    * Is the effect warming or cooling?
>    * Which type of cloud has the bigger effect on $I_{\text{out}}$?
>    * Why do you see the difference between the two types of clouds?

### Answer 1(a)

```{r answer_1a}
#| echo=FALSE, message=FALSE, warning=FALSE
clear_sky_file <- file.path(data_dir, "ch_07", "ex_7_1_a_clear.txt")
if (file.exists(clear_sky_file)) {
  modtran_clear <- read_modtran(clear_sky_file)
} else {
  modtran_clear <- run_modtran(clear_sky_file, atmosphere = "tropical")
}

alto_file <- file.path(data_dir, "ch_07", "ex_7_1_a_alto.txt")
if (file.exists(alto_file)) {
  modtran_alto <- read_modtran(alto_file)
} else {
  modtran_alto <- run_modtran(alto_file, atmosphere = "tropical",
                              clouds = "altostratus")
}

stratus_file <- file.path(data_dir, "ch_07", "ex_7_1_a_stratus.txt")
if (file.exists(stratus_file)) {
  modtran_stratus <- read_modtran(stratus_file)
} else {
  modtran_stratus <- run_modtran(stratus_file, atmosphere = "tropical",
                              clouds = "stratus")
}

i_out_clear <- modtran_clear$i_out
i_out_alto <- modtran_alto$i_out
i_out_stratus <- modtran_stratus$i_out

alto_effect <- i_out_alto - i_out_clear
stratus_effect <- i_out_stratus - i_out_clear
```

With clear skies,
$I_{\text{out}} = `r format_md(i_out_clear, digits = 2)`~\text{W}/\text{m}^2$.
With altostratus clouds,
$I_{\text{out}} = `r format_md(i_out_alto, digits = 2)`~\text{W}/\text{m}^2$,
which is $`r format_md(abs(alto_effect), digits = 2)`~~\text{W}/\text{m}^2$
`r ifelse(alto_effect >= 0, "greater", "less")`, so altostratus clouds
have a `r ifelse(alto_effect >= 0, "cooling", "warming")` effect.

With stratus clouds, which are closer to the ground,
$I_{\text{out}} = `r format_md(i_out_stratus, digits = 2)`~\text{W}/\text{m}^2$,
which is $`r format_md(abs(stratus_effect), digits = 2)`~~\text{W}/\text{m}^2$
`r ifelse(stratus_effect >= 0, "greater", "less")`, so stratus clouds
have a `r ifelse(stratus_effect >= 0, "cooling", "warming")` effect.

`r ifelse(abs(alto_effect) > abs(stratus_effect), "Altostratus", "Stratus")`
clouds have a bigger effect on $I_{\text{out}}$. This is because
altostratus clouds are higher in the atmosphere
(the tops are at 3 km, compared to 1 km for stratus), so the tops of
altostratus clouds are cooler than the tops of stratus clouds, and thus
emits less longwave radiation.

> b) Starting set the altitude to zero and select "Looking up." When you
>    are looking up, the model reports the longwave radiation coming down to
>    the surface from the atmosphere and hitting the earth's surface. This
>    heat is in addition to whatever heat the earth gets from shortwave solar
>    radiation.
>
>    For simplicity, think of this as the conditions at night, when the sun
>    is not shining: without sunlight, the temperature of the ground will be
>    determined by balancing the outgoing heat with the heat radiated downward
>    by the warm atmosphere and clouds.
>
>    * First, note the downward longwave heat flux ($I_{\text{down}}$) with
>      clear sky (no clouds or rain).
>    * Then turn on altostratus clouds and note the change in
>      $I_{\text{down}}$.
>    * Next, turn on stratus clouds and note the change in $I_{\text{down}}$.
>
>    Answer the following questions:
>
>    * How does \(I_{\text{down}}\) change when you add clouds?
>    * Do the clouds have a heating or cooling effect? Why?
>    * Which clouds have a greater heating or cooling effect? Why?

### Answer 1(b)

```{r answer_1b}
#| echo=FALSE, warning=FALSE, message=FALSE
clear_sky_down_file <- file.path(data_dir, "ch_07", "ex_7_1_b_clear.txt")
if (file.exists(clear_sky_down_file)) {
  modtran_clear_down <- read_modtran(clear_sky_down_file)
} else {
  modtran_clear_down <- run_modtran(clear_sky_down_file, atmosphere = "tropical",
                               altitude_km = 0, looking = "down")
}

alto_down_file <- file.path(data_dir, "ch_07", "ex_7_1_b_alto.txt")
if (file.exists(alto_down_file)) {
  modtran_alto_down <- read_modtran(alto_down_file)
} else {
  modtran_alto_down <- run_modtran(alto_down_file, atmosphere = "tropical",
                                   clouds = "altostratus",
                                   altitude_km = 0, looking = "down")
}

stratus_down_file <- file.path(data_dir, "ch_07", "ex_7_1_b_stratus.txt")
if (file.exists(stratus_down_file)) {
  modtran_stratus_down <- read_modtran(stratus_down_file)
} else {
  modtran_stratus_down <- run_modtran(stratus_down_file, atmosphere = "tropical",
                              clouds = "stratus",
                                   altitude_km = 0, looking = "down")
}

i_down_clear <- modtran_clear_down$i_out
i_down_alto <- modtran_alto_down$i_out
i_down_stratus <- modtran_stratus_down$i_out

alto_effect <- i_down_alto - i_down_clear
stratus_effect <- i_down_stratus - i_down_clear

```

With clear skies,
$I_{\text{down}} = `r format_md(i_down_clear, digits = 2)`~\text{W}/\text{m}^2$.
With altostratus clouds,
$I_{\text{down}} = `r format_md(i_down_alto, digits = 2)`~\text{W}/\text{m}^2$,
which is $`r format_md(abs(alto_effect), digits = 2)`~~\text{W}/\text{m}^2$
`r ifelse(alto_effect >= 0, "greater", "less")`, so altostratus clouds
have a `r ifelse(alto_effect >= 0, "warming", "cooling")` effect.

With stratus clouds, which are closer to the ground,
$I_{\text{down}} = `r format_md(i_down_stratus, digits = 2)`~\text{W}/\text{m}^2$,
which is $`r format_md(abs(stratus_effect), digits = 2)`~~\text{W}/\text{m}^2$
`r ifelse(stratus_effect >= 0, "greater", "less")`, so stratus clouds
have a `r ifelse(stratus_effect >= 0, "warming", "cooling")` effect.

`r ifelse(abs(alto_effect) > abs(stratus_effect), "Altostratus", "Stratus")`
clouds have a bigger effect on $I_{\text{down}}$. This is because
altostratus clouds are higher in the atmosphere
(the bottoms are at 2.4 km, compared to 0.33 km for stratus), so the bottoms of
altostratus clouds are cooler than the bottoms of stratus clouds, and thus
emits less longwave radiation.


## **Exercise #2: Cloud Effects with RRTM**

> Use the RRTM model.
>
> a) First set the model to its default parameters. It should report,
>    "If the earth has these properties ... then it loses as much energy as
>    it gains."
>    Move your mouse over the arrows at the top and bottom of the graph.
>    The orange arrows are shortwave (mostly visible) light, and the purple
>    arrows are longwave (far-infrared) radiation. How much shortwave and
>    longwave light is absorbed by the ground (the downward arrows at the
>    bottom) and how much of each is emitted to space (the upward arrows at
>    the top)?

### Answer for 2(a)

```{r answer_2a}
#| echo=FALSE, warning=FALSE, message=FALSE
rrtm_clear_file <- file.path(data_dir, "ch_07", "ex_7_2_a_clear.txt")
if (file.exists(rrtm_clear_file)) {
  rrtm_clear <- read_rrtm(rrtm_clear_file)
} else {
  rrtm_clear <- run_rrtm(rrtm_clear_file)
}

bottom <- rrtm_clear$fluxes %>% filter(altitude == 0)
top <- rrtm_clear$fluxes %>% top_n(n = 1, wt = altitude)

lw_down <- bottom$lw_down
sw_down <- bottom$sw_down
lw_up <- top$lw_up
sw_up <- top$sw_up
```

At the bottom, the intensity of longwave radiation is `r lw_down` W/m^2^
and the intensity of shortwave radiation is `r sw_down` W/m^2^.

At the top of the atmosphere, the intensity of longwave radiation is
`r lw_up` W/m^2^
and the intensity of shortwave radiation is `r sw_up` W/m^2^.

> b) Next, add 100% high clouds (set "High cloud (fraction)" to 1.0).
>    Record the total gain or loss of of heat, and the amount of shortwave
>    and longwave radiation absorbed by the surface and emitted to space.
>    Clouds affect both longwave and shortwave heat fluxes. Which kind of
>    radiation changed more?

### Answer for 2(b)

```{r answer_2b}
#| echo=FALSE, warning=FALSE, message=FALSE
rrtm_high_cloud_file <- file.path(data_dir, "ch_07", "ex_7_2_b.txt")
if (file.exists(rrtm_high_cloud_file)) {
  rrtm_high_cloud <- read_rrtm(rrtm_high_cloud_file)
} else {
  rrtm_high_cloud <- run_rrtm(rrtm_high_cloud_file, high_cloud_frac = 1.0)
}

bottom_hc <- rrtm_high_cloud$fluxes %>% filter(altitude == 0)
top_hc <- rrtm_high_cloud$fluxes %>% top_n(n = 1, wt = altitude)

lw_down_hc <- bottom_hc$lw_down
sw_down_hc <- bottom_hc$sw_down
lw_up_hc <- top_hc$lw_up
sw_up_hc <- top_hc$sw_up

delta_lw_down_hc <- lw_down_hc - lw_down
delta_sw_down_hc <- sw_down_hc - sw_down
delta_lw_up_hc <- lw_up_hc - lw_up
delta_sw_up_hc <- sw_up_hc - sw_up

df <- tibble(
  altitude = c("Ground", "Ground", "Top of atmosphere", "Top of atmosphere"),
  wavelength = c("Shortwave", "Longwave", "Shortwave", "Longwave"),
  direction = c("Down", "Down", "Up", "Up"),
  clear = c(sw_down, lw_down, sw_up, lw_up),
  cloud = c(sw_down_hc, lw_down_hc, sw_up_hc, lw_up_hc),
  change = c(delta_sw_down_hc, delta_lw_down_hc, delta_sw_up_hc,
             delta_lw_up_hc)
)

knitr::kable(df, col.names = c("Altitude", "Wavelength", "Direction",
                               "Clear Sky", "High Clouds", "Difference"))
```

The biggest change was in outgoing longwave radiation from the top of the
atmosphere, which dropped by `r abs(delta_lw_up_hc)` W/m^2^.

> c) Now do the same thing for low clouds: set "High cloud (fraction)" to
>    zero and "Low cloud (fraction)" to 1.0. Which kind of radiation changed
>    more, compared to the no-cloud condition?

### Answer for 2(c)

```{r answer_2c}
#| echo=FALSE, warning=FALSE, message=FALSE
rrtm_low_cloud_file <- file.path(data_dir, "ch_07", "ex_7_2_c.txt")
if (file.exists(rrtm_low_cloud_file)) {
  rrtm_low_cloud <- read_rrtm(rrtm_low_cloud_file)
} else {
  rrtm_low_cloud <- run_rrtm(rrtm_low_cloud_file, low_cloud_frac = 1.0)
}

bottom_lc <- rrtm_low_cloud$fluxes %>% filter(altitude == 0)
top_lc <- rrtm_low_cloud$fluxes %>% top_n(n = 1, wt = altitude)

lw_down_lc <- bottom_lc$lw_down
sw_down_lc <- bottom_lc$sw_down
lw_up_lc <- top_lc$lw_up
sw_up_lc <- top_lc$sw_up

delta_lw_down_lc <- lw_down_lc - lw_down
delta_sw_down_lc <- sw_down_lc - sw_down
delta_lw_up_lc <- lw_up_lc - lw_up
delta_sw_up_lc <- sw_up_lc - sw_up

df <- tibble(
  altitude = c("Ground", "Ground", "Top of atmosphere", "Top of atmosphere"),
  wavelength = c("Shortwave", "Longwave", "Shortwave", "Longwave"),
  direction = c("Down", "Down", "Up", "Up"),
  clear = c(sw_down, lw_down, sw_up, lw_up),
  cloud = c(sw_down_lc, lw_down_lc, sw_up_lc, lw_up_lc),
  change = c(delta_sw_down_lc, delta_lw_down_lc, delta_sw_up_lc,
             delta_lw_up_lc)
)

knitr::kable(df, col.names = c("Altitude", "Wavelength", "Direction",
                               "Clear Sky", "Low Clouds", "Difference"))
```

The biggest change was in the longwave coming down to the ground, which
increased by `r delta_lw_down_lc` W/m^2^.


> d) With the low cloud fraction still set to 1.0 (100%), change the drop
>    radius from 10 to 8 $\mu$m. How does this change the heat flux?

### Answer for 2(d)

```{r answer_2d}
#| echo=FALSE, warning=FALSE, message=FALSE
rrtm_low_cloud_small_drop_file <- file.path(data_dir, "ch_07", "ex_7_2_d.txt")
if (file.exists(rrtm_low_cloud_small_drop_file)) {
  rrtm_low_cloud_small_drop <- read_rrtm(rrtm_low_cloud_small_drop_file)
} else {
  rrtm_low_cloud_small_drop <- run_rrtm(rrtm_low_cloud_small_drop_file,
                                        low_cloud_frac = 1.0,
                                        cloud_drop_radius = 8)
}

bottom_lc_8 <- rrtm_low_cloud_small_drop$fluxes %>% filter(altitude == 0)
top_lc_8 <- rrtm_low_cloud_small_drop$fluxes %>% top_n(n = 1, wt = altitude)

lw_down_lc_8 <- bottom_lc_8$lw_down
sw_down_lc_8 <- bottom_lc_8$sw_down
lw_up_lc_8 <- top_lc_8$lw_up
sw_up_lc_8 <- top_lc_8$sw_up

delta_lw_down_lc_8 <- lw_down_lc_8 - lw_down_lc
delta_sw_down_lc_8 <- sw_down_lc_8 - sw_down_lc
delta_lw_up_lc_8 <- lw_up_lc_8 - lw_up_lc
delta_sw_up_lc_8 <- sw_up_lc_8 - sw_up_lc

i_out_large <- rrtm_low_cloud$i_out
i_in_large <- rrtm_low_cloud$i_in
i_out_small <- rrtm_low_cloud_small_drop$i_out
i_in_small <- rrtm_low_cloud_small_drop$i_in

df <- tibble(
  altitude = c("Ground", "Ground", "Top of atmosphere", "Top of atmosphere"),
  wavelength = c("Shortwave", "Longwave", "Shortwave", "Longwave"),
  direction = c("Down", "Down", "Up", "Up"),
  large = c(sw_down_lc, lw_down_lc, sw_up_lc, lw_up_lc),
  small = c(sw_down_lc_8, lw_down_lc_8, sw_up_lc_8, lw_up_lc_8),
  change = c(delta_sw_down_lc_8, delta_lw_down_lc_8, delta_sw_up_lc_8,
             delta_lw_up_lc_8)
)

knitr::kable(df, col.names = c("Altitude", "Wavelength", "Direction",
                               "Large drops", "Small drops", "Difference"))
```

There is very little change in the longwave, but big reduction in shortwave
reaching the ground and a big increase in shortwave reflected back to space.

In terms of total heat fluxes, with the 10 $\mu$m drops, we had
$I_{\text{in}} = `r i_in_large`$ and $I_{\text{out}} = `r i_out_large`$.

For the 8 $\mu$m drops, we had
$I_{\text{in}} = `r i_in_small`$ and $I_{\text{out}} = `r i_out_small`$.


> e) Now set the cloud fraction to zero and double the CO~2~. How does this
>    change the heat flux? How does the effect of doubling CO~2~ compare to
>    the effect of changing the droplet size for the low clouds? You can see
>    how important it is to get the cloud droplet size right in climate
>    models!

### Answer for 2(e)

```{r answer_2e}
#| echo=FALSE, warning=FALSE, message=FALSE
rrtm_co2_file <- file.path(data_dir, "ch_07", "ex_7_2_e.txt")
if (file.exists(rrtm_co2_file)) {
  rrtm_co2 <- read_rrtm(rrtm_co2_file)
} else {
  rrtm_co2 <- run_rrtm(rrtm_co2_file, co2_ppm = 800)
}

bottom_co2 <- rrtm_co2$fluxes %>% filter(altitude == 0)
top_co2 <- rrtm_co2$fluxes %>% top_n(n = 1, wt = altitude)

lw_down_co2 <- bottom_co2$lw_down
sw_down_co2 <- bottom_co2$sw_down
lw_up_co2 <- top_co2$lw_up
sw_up_co2 <- top_co2$sw_up

delta_lw_down_co2 <- lw_down_co2 - lw_down
delta_sw_down_co2 <- sw_down_co2 - sw_down
delta_lw_up_co2 <- lw_up_co2 - lw_up
delta_sw_up_co2 <- sw_up_co2 - sw_up

i_out <- rrtm_clear$i_out
i_in <- rrtm_clear$i_in
i_out_co2 <- rrtm_co2$i_out
i_in_co2 <- rrtm_co2$i_in


df <- tibble(
  altitude = c("Ground", "Ground", "Top of atmosphere", "Top of atmosphere"),
  wavelength = c("Shortwave", "Longwave", "Shortwave", "Longwave"),
  direction = c("Down", "Down", "Up", "Up"),
  clear = c(sw_down, lw_down, sw_up, lw_up),
  cloud = c(sw_down_co2, lw_down_co2, sw_up_co2, lw_up_co2),
  change = c(delta_sw_down_co2, delta_lw_down_co2, delta_sw_up_co2,
             delta_lw_up_co2)
)

knitr::kable(df, col.names = c("Altitude", "Wavelength", "Direction",
                               "Normal CO2", "Double CO2", "Difference"))
```

In terms of total heat fluxes, with 400 ppm CO~2~, we had
$I_{\text{in}} = `r i_in`$ and $I_{\text{out}} = `r i_out`$.

For the 800 ppm CO~2~, we had
$I_{\text{in}} = `r i_in_co2`$ and $I_{\text{out}} = `r i_out_co2`$.

Changing the size of cloud drops affects the heat balance about half as much
as doubling CO~2~,
but CO~2~ mostly affects $I_{\text{out}}$ and droplet size mostly
affect $I_{\text{in}}$


## **Exercise #3: Water-Vapor Feedback**

> The "climate sensitivity" ($\Delta T_{\text{2x}}$) refers to the change in
> temperature when you double the amount of CO~2~ in the atmosphere. Here,
> we will examine how the water vapor feedback affects climate sensitivity.
> 
> a) Run the RRTM model with the default parameters. Write down the ground
>    temperature.
>    
>    Next, double CO~2~ and note the change in energy balance.
>    At the beginning of the exercise, the earth was in radiative equilibrium.
>    Changing CO~2~ disturbed this equilibrium and produced a radiative
>    imbalance where $I_{\text{out}} \neq I_{\text{in}}$.
>    Adjust the surface temperature to bring the earth back into balance.
>    
>    * What is the new temperature?
>    * How much did the earth warm or cool? This temperature change for
>      doubling CO~2~ is what we call the **climate sensitivity**,
>      or $\Delta T_{2\times\COO}$.

### Answer for 3(a)

```{r answer_3_a}
#| echo=FALSE, warning=FALSE, message=FALSE
t0 <- 284.42
t_2x <- 286.9

delta_2x <- t_2x - t0
```
When I run the RRTM model with the default parameters, the ground 
temperature is `r t0` K. 

When I double CO~2~, initially, this reduces $I_{\text{out}}$ so the Earth
gains energy at a rate of 4.2 W/m^2^. I adjust the surface temperature by 
trial and error, raising it when the Earth gains energy and reducing it when
Earth loses energy, until I find that at `r t_2x` K, the Earth loses as much
energy as it gains. 

The earth warmed up by `r delta_2x` K when we doubled CO~2~, so this is the
_climate sensitivity_.

> b) Set CO~2~ back to the default value (400 ppm) and set relative humidity
>    to zero. This turns off the water vapor feedback.
>    Adjust the surface temperature offset until the earth loses as much
>    energy as it gains.
>    Write down this temperature.
>    
>    Now double the CO~2~ and adjust the surface temperature to bring the
>    heat back into balance.
>    
>    * How much did the temperature change?

### Answer for 3(b)

```{r answer_3_b}
#| echo=FALSE, warning=FALSE, message=FALSE
t0_rh <- 261.55
t_2x_rh <- 262.65
delta_2x_rh <- t_2x_rh - t0_rh
```

When I return RRTM to its default conditions and turn off the water vapor
feedback by setting relative humidity to zero, 
this increases $I_{\text{out}}$ so the Earth loses energy at a rate of 
91.1 W/m^2^. I adjust the surface temperature by trial and error until I find 
that at `r t0_rh` K, the Earth loses as much energy as it gains. 

Then I double CO~2~ to 800 ppm and the Earth gains energy at a rate of
3.8 W/m^2^. Again, I adjust the surface temperature by trial and error until
I find that at `r t_2x_rh`, the Earth loses as much energy as it gains.

The climate sensitivity for 0% relative humidity is `r delta_2x_rh`.

> c) Compare the climate sensitivity with zero relative humidity to the
>    sensitivity with the default value of 80% relative humidity. The
>    difference is the effect of water vapor feedback.
>    
>    * What was the amplification factor of the water vapor feedback
>      (the ratio of the climate sensitivity $\Delta T_{2\times\COO}$ with
>      water vapor feedback to $\Delta T_{2\times\COO}$ without it)?

### Answer for 3(c)

With relative humidity set to zero (no water-vapor feedback), the climate
sensitivity is `r delta_2x_rh`. With relative humidity set to 80% 
(normal water-vapor feedback) the climate sensitivity is `r delta_2x`.
The amplification factor for water-vapor feedback is the ratio of
these numbers

$$\text{amplification factor} = \frac{`r delta_2x` K}{`r delta_2x_rh` K} = 
   `r format_md(delta_2x / delta_2x_rh, digits = 2)`$$
   
Water vapor feedback more than doubles the climate sensitivity!
