---
title: Chapter 4 Homework Answers
hw_number: 2
pubdate: '2023-02-06'
date: '2023-02-01'
pdf_url: /files/homework_solutions/hw_02/homework_02_answers.pdf
slug: homework_02_answers
output:
  blogdown::html_page:
    md_extensions: +tex_math_single_backslash+compact_definition_lists
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 3.0
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

library(tidyverse)
library(lubridate)

data_dir = "_data"
script_dir = "_scripts"

# Create a data directory if one does not exist.
if (!dir.exists(data_dir)) dir.create(data_dir)
if (!dir.exists(file.path(data_dir, "ch_04")))
    dir.create(file.path(data_dir, "ch_04"))

source(file.path(script_dir, 'utils.R'), chdir = T)
source(file.path(script_dir, 'format_md.R'), chdir = T)
source(file.path(script_dir, 'modtran.R'), chdir = T)

epsilon = 1
alpha = 0.30
sigma = 5.67E-8
I_solar = 1350
```

# Chapter 4: Longwave Radiation Spectrum

## Exercise 4.1: Methane

> Methane has a current concentration of 1.7 ppm in the atmosphere and
> is doubling at a faster rate than CO~2~.

### Part (a)

> **Would an additional 10 ppm of methane in the atmosphere have a larger or
> smaller impact on the outgoing IR flux than an additional 10 ppm of CO~2~
> at current concentrations?**

```{r ex_4_1_a, include=TRUE, message=FALSE, warning=FALSE}
base_co2 = 400 # parts per million
base_ch4 = 1.7 # parts per million

baseline_file <- file.path(data_dir, "ch_04", "ex_4_1_baseline.txt")
if (file.exists(baseline_file)) {
  modtran_baseline = read_modtran(baseline_file)
} else {
  modtran_baseline = run_modtran(baseline_file, atmosphere = "tropical",
                                 co2_ppm = base_co2,
                                 ch4_ppm = base_ch4)
}

co2_file <- file.path(data_dir, "ch_04", "ex_4_1_plus_10_co2.txt")
if (file.exists(co2_file)) {
  modtran_plus_10_co2 = read_modtran(co2_file)
} else {
  modtran_plus_10_co2 = run_modtran(co2_file, atmosphere = "tropical",
                                    co2_ppm = base_co2 + 10,
                                    ch4_ppm = base_ch4)
}

ch4_file <- file.path(data_dir, "ch_04", "ex_4_1_plus_10_ch4.txt")
if (file.exists(ch4_file)) {
  modtran_plus_10_ch4 = read_modtran(ch4_file)
} else {
  modtran_plus_10_ch4 = run_modtran(ch4_file, atmosphere = "tropical",
                                    co2_ppm = base_co2,
                                    ch4_ppm = base_ch4 + 10)
}

i_out_baseline = modtran_baseline$i_out
i_out_co2 = modtran_plus_10_co2$i_out
i_out_ch4 = modtran_plus_10_ch4$i_out
delta_i_co2 = i_out_co2 - i_out_baseline
delta_i_ch4 = i_out_ch4 - i_out_baseline
```

**Answer:**
I ran MODTRAN three times.

* One run was a baseline, which used the current concentrations of CO~2~ and
  CH~4~.
* Then I ran MODTRAN with the CO~2~ concentration increased by 10 ppm.
* Finally, I ran it a third time with the baseline value for CO~2~, but with
  CH~4~ increased by 10 ppm.

For the baseline run, the intensity of outgoing longwave light
was `r format_md(i_out_baseline, digits = 3)` Watts per square meter.
Increasing CO~2~ by 10 ppm decreased the outgoing longwave light by
`r format_md(-delta_i_co2, digits = 1)` W/m^2^, to
`r format_md(i_out_co2, digits = 3)`
and increasing CH~4~ by
10 ppm decreased the outgoing longwave light by
`r format_md(-delta_i_ch4, digits = 2)` W/m^2^ to
`r format_md(i_out_ch4, digits = 3)`.
The change in I~out~ for methane is is around
`r format_md(delta_i_ch4 / delta_i_co2, digits = 1)` times as much as for CO~2~.

The difference is because absorption for CO~2~ is strongly saturated, but the
absorption for CH~4~ is not saturated. Another way to think about this is that
a 10 ppm increase in CO~2~ increases the amount of CO~2~ by
`r format_md(100 * 10 / base_co2, digits = 1)`% and a 10 ppm increase in
CH~4~ increases the amount of CH~4~ by
`r format_md(100 * 10 / base_ch4, digits = 1)`%.

### Part (b)

> **Where in the spectrum does methane absorb? What concentration does it
> take to begin to saturate the absorption in this band? Explain what you are
> looking at to judge when the gas is saturated.**

**Answer:**
The plots below show the MODTRAN spectrum with all gases set to zero except
methane. Methane absorbs most strongly around 1300 wavenumbers.


```{r ex_4_1_b, include=TRUE, message=FALSE, fig.height=4, fig.width=5}
ch4_values = c(0, 2^seq(0,7))
sat_data = tibble()
for (ch4 in ch4_values) {
   filename = file.path(data_dir, "ch_04", str_c("ex_4_1_ch4_", ch4, ".txt"))
   if (file.exists(filename)) {
      mod_data = read_modtran(filename)
   } else {
      mod_data = run_modtran(filename, atmosphere = "tropical",
                             co2_ppm = 0, ch4_ppm = ch4,
                             trop_o3_ppb = 0, strat_o3_scale = 0,
                             h2o_scale = 0, freon_scale = 0,
                             altitude_km = 70)
   }
   p = plot_modtran(mod_data, descr = str_c(ch4, " ppm methane"))
   plot(p)
   if (knitr::is_latex_output()) {
     cat("\n\n")
   }
}
```


As we increase the methane concentration, the big spike around 1300 wavenumbers
gets bigger until it bottoms out on the purple line. This happens somewhere
around 8, 16, or 32 ppm, so any of those anwers would be correct. But the
spectrum is complicated and so is its saturation, so other answers are plausible
if they are supported by sound reasoning.

### Part (c)

> **Would a doubling of methane have as great an impact on the heat balance as
> a doubling of CO~2~?**

```{r ex_4_1_c, include=TRUE, message=FALSE}
base_co2_ppm = 400
base_methane_ppm = 1.7

baseline =   read_modtran(file.path(data_dir, "ch_04", "ex_4_1_baseline.txt"))

double_co2_file <- file.path(data_dir, "ch_04", "ex_4_1_2x_co2.txt")
if (file.exists(double_co2_file)) {
  double_co2 = read_modtran(double_co2_file)
} else {
  double_co2 = run_modtran(double_co2_file, atmosphere = "tropical",
                           co2_ppm = 2 * base_co2_ppm,
                           ch4_ppm = base_methane_ppm)
}

double_ch4_file <- file.path(data_dir, "ch_04", "ex_4_1_2x_ch4.txt")
if (file.exists(double_ch4_file)) {
  double_ch4 = read_modtran(double_ch4_file)
} else {
  double_ch4 = run_modtran(double_ch4_file, atmosphere = "tropical",
                           co2_ppm = base_co2_ppm,
                           ch4_ppm = 2 * base_methane_ppm)
}

i_out_baseline = baseline$i_out
i_out_co2 = double_co2$i_out
i_out_ch4 = double_ch4$i_out

delta_i_out_co2 = i_out_co2 - i_out_baseline
delta_i_out_ch4 = i_out_ch4 - i_out_baseline
```

**Answer:**
The baseline value for I~out~ is
`r format_md(i_out_baseline, digits = 3)` W/m^2^.
If we double CO~2~, it drops to
`r format_md(i_out_co2, digits = 3)` W/m^2^, a decrease of
`r format_md(abs(delta_i_out_co2), digits = 1)` W/m^2^,
and if we double CH~4~, it drops to
`r format_md(i_out_ch4, digits = 3)` W/m^2^, a decrease of
`r format_md(abs(delta_i_out_ch4), digits = 1)` W/m^2^.
Doubling CO~2~ has the larger effect because there is a lot more CO~2~ in the
atmosphere and that is more important than how saturated its absorption is.

You can see this if you look at the plots below. Notice that the effect of
doubling CO~2~ isn't to make the big CO~2~ absorption feature get deeper, but
to make it wider. You can see this if you compare the baseline to the doubled
CO~2~ spectrum where the spectrum crosses the 240 K blackbody curve
near 600 cm^-1^ and around 750 cm^-1^.  Compare this to the very
small change in the methane spike near 1300 cm^-1^ (you have to look
very carefully at the doubled methane spectrum to notice this).

```{r ex_4_1_c_baseline_spectrum}
plot_modtran(baseline, descr = "Baseline spectrum")
```
```{r ex_4_1_c_double_co2_spectrum}
plot_modtran(double_co2, descr = "Doubled CO2")
```
```{r ex_4_1_c_double_ch4_spectrum}
plot_modtran(double_ch4, descr = "Doubled CH4")
```

### Part (d)

> **What is the "equivalent CO~2~" of doubling atmospheric methane? That is to
> say, how many ppm of CO~2~ would lead to the same change in outgoing IR
> radiation energy flux as doubling methane? What is the ratio of ppm CO~2~
> change to ppm methane change?**

```{r ex_4_1_d, include=TRUE, message=FALSE}
delta_co2 = 80
delta_ch4 = 1.7

co2_match_file <- file.path(data_dir, "ch_04", "ex_4_1_co2_match.txt")
modtran_match_co2 = run_modtran(co2_match_file, atmosphere = "tropical",
                                co2_ppm = 400 + delta_co2)
```

**Answer:**
When we double CH~4~, I~out~ is
`r format_md(i_out_ch4, digits = 4)` W/m^2^.
We need to adjust CO~2~ to produce the same I~out~ with the defailt value of
1.7 ppm CH~4~. After some trial and error, this turns out to be about
`r format(400 + delta_co2, digits = 1)` ppm,
which has I~out~ = `r format_md(modtran_match_co2$i_out, digits = 4)`.
The ratio of $\Delta\COO$ to $\Delta\methane$ is
$`r delta_co2`/`r delta_ch4` = `r format_md(delta_co2/delta_ch4, digits = 2, output_format="bare-latex")`$

## Exercise 4.3: Water vapor

> Our theory of climate presumes that an increase in the temperature at ground
> level will lead to an increase in the outgoing IR energy flux at the top of
> the atmosphere.

### Part (a)

> **How much extra outgoing IR would you get by raising the temperature of the
> ground by 5&deg;C? What effect does the ground temperature have on the
> shape of the outgoing IR spectrum and why?**

```{r ex_4_3_a, include=TRUE, message=FALSE}

modtran_baseline_file <- file.path(data_dir, "ch_04", "ex_4_3_baseline.txt")
if (file.exists(modtran_baseline_file)) {
  modtran_baseline = read_modtran(modtran_baseline_file)
} else {
  modtran_baseline = run_modtran(modtran_baseline_file, 
                                 atmosphere = "tropical")
}

modtran_plus_5_file <- file.path(data_dir, "ch_04", "ex_4_3_t_plus_5")
if (file.exists(modtran_plus_5_file)) {
  modtran_plus_5 = read_modtran(modtran_plus_5_file)
} else {
  modtran_plus_5 = run_modtran(modtran_plus_5_file, atmosphere = "tropical",
                               delta_t = 5)
}

i_out_baseline = modtran_baseline$i_out
i_out_plus_5 = modtran_plus_5$i_out

p_baseline = plot_modtran(modtran_baseline, descr = "Baseline spectrum")
p_plus_5 = plot_modtran(modtran_plus_5,
                        descr = "Ground temperature raised 5K")
plot(p_baseline)
plot(p_plus_5)
```

**Answer:**
Raising the ground temperature raises the entire spectrum, and the intensity of
outgoing longwave radiation increases from 
`r format_md(i_out_baseline, digits = 4)` W/m^2^ to
`r format_md(i_out_plus_5, digits = 4)` W/m^2^: a change of 
`r format_md(i_out_plus_5 - i_out_baseline, digits = 3)` W/m^2^.

### Part (b)

> **More water can evaporate into warm air than into cool air. Change the
> model settings to hold the water vapor at constant relative humidity
> rather than constant vapor pressure (the default), calculate the change
> in outgoing IR energy flux for a 5&deg;C temperature increase.
> Is it higher or lower? Does water vapor make the Earth more sensitive to
> CO~2~ increases or less sensitive?**

```{r ex_4_3_b, include=TRUE, message=FALSE}
modtran_vp_file <- file.path(data_dir, "ch_04", "ex_4_3_vap_press.txt")
if (file.exists(modtran_vp_file)) {
  modtran_vp = read_modtran(modtran_vp_file)
} else {
  modtran_vp = run_modtran(modtran_vp_file, atmosphere = "tropical",
                           delta_t = 5, h2o_fixed = "vapor pressure")
}

modtran_rh_file <- file.path(data_dir, "ch_04", "ex_4_3_rel_hum.txt")
if (file.exists(modtran_rh_file)) {
  modtran_rh = read_modtran(modtran_rh_file)
} else {
  modtran_rh = run_modtran(modtran_rh_file, atmosphere = "tropical",
                           delta_t = 5, h2o_fixed = "relative humidity")
}

# i_out_baseline = modtran_baseline$i_out
i_out_vp = modtran_vp$i_out
i_out_rh = modtran_rh$i_out

p_vp = plot_modtran(modtran_vp, descr = "Constant vapor pressure",
                    i_out_ref = i_out_baseline)
p_rh = plot_modtran(modtran_rh, descr = "Constant relative humidity",
                    i_out_ref = i_out_baseline)
plot(p_vp)
plot(p_rh)
```

**Answer:**
Raising the ground temperature with constant relative humidity 
raises the intensity of outgoing longwave radiation from 
`r format_md(i_out_baseline, digits = 4)` W/m^2^ to
`r format_md(i_out_rh, digits = 4)` W/m^2^: a change of 
`r format_md(i_out_rh - i_out_baseline, digits = 3)` W/m^2^.

This is a smaller change than when we raised the surface temperature with
constant vapor pressure. 

This means that compensating for a change in CO~2~ would require a bigger 
change in temperature with fixed relative humidity, so the climate is 
**more sensitive** to changes in CO~2~ when relative humidity is fixed.

### Part (c)

```{r ex_4_3_c_setup, eval=TRUE, results="hide"}
modtran_baseline = read_modtran(file.path(data_dir, "ch_04", "ex_4_3_baseline.txt"))
i_baseline = modtran_baseline$i_out

doubled_co2 <- FALSE
if (doubled_co2) {
  new_co2 <- 800
  dt_vp = 0.76
  dt_rh = 1.21
  file_tag <- "2x_co2"
} else {
  new_co2 <- 550
  dt_vp = 0.345
  dt_rh = 0.55
  file_tag <- "550_co2"
}
```

> **Now see this effect in another way.**
>
> * **Starting from the default base case, record the total outgoing
>   IR flux.**
>
> * **`r ifelse(doubled_co2, "Double CO~2~", "Increase CO~2~ to 550 ppm")`. 
>   The temperature in the model stays the same 
>   (that's how the model is written), but the outgoing IR flux goes down.**
>
> * **Using constant water vapor pressure, adjust the temperature offset
>   until you get the original IR flux back again. Record the change in
>   temperature.**
>
> * **Now repeat the calculation, but holding the relative humidity fixed
>   instead of the water vapor pressure.**
>   Does the increase in CO~2~ drive a bigger or smaller teemperature change?
> 

```{r ex_4_3_c, include=TRUE, message=FALSE}
modtran_vp_file <- file.path(data_dir, "ch_04", 
                             str_c("ex_4_3", file_tag, "vp.txt"))
if (file.exists(modtran_vp_file)) {
  modtran_vp = read_modtran(modtran_vp_file)
} else {
  modtran_vp = run_modtran(modtran_vp_file, atmosphere = "tropical",
                           co2_ppm = new_co2)
}

i_vp = modtran_vp$i_out

modtran_vp_dt_file <- file.path(data_dir, "ch_04",
                                str_c("ex_4_3", file_tag, "vp_dt.txt"))
if (file.exists(modtran_vp_dt_file)) {
  modtran_vp_dt = read_modtran(modtran_vp_dt_file)
} else {
  modtran_vp_dt = run_modtran(modtran_vp_dt_file, atmosphere = "tropical",
                              co2_ppm = new_co2, delta_t = dt_vp)
}

i_vp_dt = modtran_vp_dt$i_out

modtran_rh_file <- file.path(data_dir, "ch_04", 
                             str_c("ex_4_3", file_tag, "rh.txt"))
if (file.exists(modtran_rh_file)) {
  modtran_rh = read_modtran(modtran_rh_file)
} else {
  modtran_rh = run_modtran(modtran_rh_file, atmosphere = "tropical",
                           co2_ppm = new_co2, h2o_fixed = "relative humidity")
}

i_rh = modtran_rh$i_out

modtran_rh_dt_file <- file.path(data_dir, "ch_04",
                                str_c("ex_4_3", file_tag, "rh_dt.txt"))
if (file.exists(modtran_rh_dt_file)) {
  modtran_rh_dt = read_modtran(modtran_rh_dt_file)
} else {
  modtran_rh_dt = run_modtran(modtran_rh_dt_file, atmosphere = "tropical",
                              co2_ppm = new_co2, delta_t = dt_rh, 
                              h2o_fixed = "relative humidity")
}

i_rh_dt = modtran_rh_dt$i_out

feedback_factor = dt_rh / dt_vp
```

**Answer:**
In the baseline case,
I~out~ = `r format_md(i_baseline, digits = 4)` W/m^2^.
When we double CO~2~ with constant water vapor pressure, I~out~ drops to
`r format_md(i_vp, digits = 4)` W/m^2^
and we have to raise the ground temperature by
`r format_md(dt_vp, digits = 3)` K
to bring I~out~ back to
`r format_md(i_vp_dt, digits = 4)` W/m^2^.

When we double CO~2~ with constant water relative humidity, I~out~ drops to
`r format_md(i_rh, digits = 4)` W/m^2^
and we have to raise the ground temperature by
`r format_md(dt_rh, digits = 3)` K
to bring I~out~ back to
`r format_md(i_rh_dt, digits = 4)` W/m^2^.

Allowing water vapor to respond to changes in temperature means that CO~2~
causes more warming. 

<!--
The feedback factor is the ratio of the temperature change with relative
humidity fixes to the temperature change with vapor pressure fixed:
$f = `r format_md(dt_rh, digits = 3)`~\text{K} /
`r format_md(dt_vp, digits = 3)`~\text{K} =
`r format_md(feedback_factor, digits = 2)`$.
-->

Notice that there is no difference between holding vapor pressure constant and
holding relative humidity constant until the temperature changes.

