---
title: Chapter 3 Homework Answers
hw_number: 1
pubdate: '2023-02-01'
date: '2023-01-25'
pdf_url: /files/homework_solutions/hw_01/homework_01_ch_3_answers.pdf
slug: homework_01_ch_3_answers
output:
  blogdown::html_page:
    md_extensions: +tex_math_single_backslash+compact_definition_lists
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 3.0
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(lubridate)

data_dir = "_data"
script_dir = "_scripts"

# Create a data directory if one does not exist.
if (!dir.exists(data_dir)) dir.create(data_dir)

source(file.path(script_dir, 'utils.R'), chdir = T)
source(file.path(script_dir, 'format_md.R'), chdir = T)
source(file.path(script_dir, 'modtran.R'), chdir = T)
source(file.path(script_dir, "layer_diagram.R"), chdir = TRUE)

epsilon = 1
alpha = 0.30
sigma = 5.67E-8
I_solar = 1350
```

# Chapter 3: Layer Models

For these exercises, use the following numbers:

* I~solar~ = 1350 W/m^2^
* $\sigma = 5.67 \times 10^{-8}$
* $\alpha = 0.30$
* $\varepsilon = 1.0$

```{r set_constants, include=TRUE}
I_solar = 1350
sigma = 5.67E-8
alpha = 0.30
epsilon = 1.0
```

## Exercise 2

> **A Two-Layer Model.** Insert another atmospheric layer into the model, just 
> like the first one. The layer is transparent to visible light but a blackbody 
> for infrared.

```{r two_layer_figure, fig.cap = "An energy diagram for a planet with two panes of glass for an atmosphere. The intensity of absorbed visible light is $(1 - \\alpha) I_{solar} / 4$."}
make_layer_diagram(2)
```

> a) Write the energy budgets for both atmospheric layers, for the ground, and 
> for the Earth as a whole, like we did for the One-Layer Model.

**Answer:** 

At each layer, calculate the heat in (in the diagram, this is the sum of
all the intensities with arrows that end at the atmospheric layer) and the
heat out (the sum of all the intensities that start at the atmospheric
layer and have arrows pointing away from it).

* Top of the atmosphere:
  $I_{2,\text{up}} = I_{\text{visible}} = (1 - \alpha) I_{\text{solar}} / 4$
* Layer 2:
  $I_{1,\text{up}} = I_{2,\text{up}} + I_{2,\text{down}}$
* Layer 1:
  $I_{\text{ground,up}} + I_{2,\text{down}} = I_{1,\text{up}} + I_{1,\text{down}}$
* Ground:
  $I_{\text{ground,up}} = I_{\text{visible}} + I_{1,\text{down}}$

> b) Manipulate the budget for the Earth as a whole to obtain the temperature 
> T~2~ of the top atmospheric layer, labeled Atmospheric Layer 2 in the figure 
> above. Does this part of the exercise seem familiar in any way? Does the term 
> ring any bells?

Top of the atmosphere:

$$
\begin{aligned}
I_{2,\text{up}} &= I_{\text{visible}} = \frac{(1 - \alpha) I_{\text{solar}}}{4}\\
I_{2,\text{up}} &= \varepsilon \sigma T_2^4\\
T_2 &= \sqrt[4]{\frac{I_{2,\text{up}}}{\varepsilon \sigma}} \\
    &= \sqrt[4]{\frac{(1 - \alpha) I_{\text{solar}}}{4 \varepsilon \sigma}} \\
    &= T_{\text{bare rock}}
\end{aligned}
$$

This is the same as the bare-rock temperature.

```{r ex_3_2_b}
I_visible = (1 - alpha) * I_solar / 4
I_2_up = I_visible
T_2 = (I_2_up / (epsilon * sigma))^0.25
```

**Answer:** The temperature of layer 2 is `r format_md(T_2, digits = 3)` K,
which is the same as the bare-rock temperature. In layer models, the top layer 
of the atmosphere is _always_ the bare-rock temperature.

> c) Insert the value you found for T~2~ into the energy budget for layer 2, and 
> solve for the temperature of layer 1 in terms of layer 2. How much bigger is 
> T~1~ than T~2~?

From the energy budget for Layer 2, 
I~1,up~ = I~2,up~ + I~2,down~.
The temperature of the bottom of the layer is the same as the temperature for
the top of the layer, so I~2,down~ = I~2,up~

```{r ex_3_2_c}
I_1_up = 2 * I_2_up
T_1 = (I_1_up / (epsilon * sigma))^0.25
```

$$
\begin{aligned}
I_{\text{2,down}} &= I_{\text{2,up}} \\
I_{\text{1,up}} &= I_{\text{2,up}} + I_{\text{2,down}} \\
 &= 2 I_{\text{2,up}} \\
T_1 &= \sqrt[4]{\frac{I_{\text{1,up}}}{\epsilon \sigma}} \\
  &= \sqrt[4]{\frac{2 I_{\text{2,up}}}{\epsilon \sigma}} \\
  &= \sqrt[4]{2} T_2 \\
  &= \sqrt[4]{2}\; T_{\text{bare rock}} \\
\end{aligned}
$$

**Answer:** The temperature of layer 1 is `r format_md(T_1, digits = 3)` K.
This is the same as the ground temperature in a 1-layer model.

> d) Now insert the value you found for T~1~ into the budget for atmospheric 
> layer 1 to obtain the temperature of the ground, T~ground~. Is the greenhouse 
> effect stronger or weaker because of the second layer?

From the energy budget for layer 1, 

$$
\begin{aligned}
I_{\text{ground,up}} + I_{\text{2,down}} &= I_{\text{1,up}} + I_{\text{1,down}} \\
I_{\text{ground,up}} &= I_{\text{1,up}} + I_{\text{1,down}} - I_{\text{2,down}} \\
\text{we} &\text{ know that} \\
I_{\text{1,down}} &= I_{\text{1,up}} \\
&\text{and} \\
I_{\text{2,down}} &= I_{\text{2,up}} \\
&\text{so} \\
I_{\text{ground,up}} &= 2 I_{\text{1,up}} - I_{\text{2,up}} \\
\text{we} &\text{ also know that} \\
I_{\text{1,up}} &= 2 I_{\text{2,up}} \\
&\text{so} \\
I_{\text{ground,up}} &= 4 I_{\text{2,up}} - I_{\text{2,up}} \\
&= 3 I_{\text{2,up}}
\end{aligned}
$$

```{r ex_3_2_d}
I_ground_up = 2 * I_1_up - I_2_up
T_ground = (I_ground_up / (epsilon * sigma))^0.25
```

$$
\begin{aligned}
T_{\text{ground}} &= 
  \sqrt[4]{\frac{I_{\text{ground,up}}}{\varepsilon \sigma}} \\
&= \sqrt[4]{\frac{3 I_{\text{2,up}}}{\varepsilon \sigma}} \\
  &= \sqrt[4]{3}\; T_2 \\
  &= \sqrt[4]{3}\; T_{\text{bare rock}}
\end{aligned}
$$

**Answer:** T~ground~ = `r format_md(T_ground, digits = 3)` K = $\displaystyle\sqrt[4]{3 T_{\text{bare rock}}}$

In a 1-layer model, the ground temperature was $\sqrt[4]{2}$ times the bare-rock 
temperature, and in a 2-layer model, the ground temperature is $\sqrt[4]{3}$ times
the bare-rock temperature.

## Exercise 3.3

```{r nuclear_winter_diagram, fig.cap="An energy diagram for a planet with an opaque pane of glass for an atmosphere. The intensity of absorbed visible light is $(1 - \\alpha) I_{solar} / 4$."}
make_nuclear_winter_diagram()
```

> **Nuclear Winter.** Let us go back to the One-Layer Model but change it so 
> that the atmospheric layer absorbs visible light rather than allowing it to 
> pass through (See the figure above). This could happen if the upper atmosphere 
> were filled with dust. For simplicity, assume that the albedo of the Earth 
> remains the same, even though in the real world it might change with a dusty 
> atmosphere.>  What is the temperature of the ground in this case?

**Answer:** 
Here, the key difference is that the heat from the sun is absorbed by the 
atmosphere instead of passing through the atmosphere to the ground.

The equation for the atmosphere is the same as in the 1-layer model because
we use the energy balance at the boundary to space:

$$I_{\text{atm, up}} = I_{\text{visible}} = \frac{(1 - \alpha) I_{\text{solar}}}{4}$$

and the temperature of the atmosphere is the bare-rock temperature, just as 
the top layer of the atmosphere is for every layer model.

However, things are different at the ground. The energy balance at the 
dusty atmosphere is

$$
\begin{aligned}
I_{\text{visible}} + I_{\text{ground,up}} &= I_{\text{atm,up}} + I_{\text{atm,down}} \\
I_{\text{ground,up}} &= I_{\text{atm,up}} + I_{\text{atm,down}} - I_{\text{visible}} \\
& \text{But} \\
I_{\text{atm,up}} &= I_{\text{atm,down}} = I_{\text{visible}}. \\
& \text{So} \\
I_{\text{ground,up}} &= I_{\text{atm,up}}. \\
\text{This} & \text{ means that} \\
T_{\text{ground}} &= T_{\text{atmosphere}} = T_{\text{bare-rock}}
\end{aligned}
$$


```{r ex_3}
I_visible = (1 - alpha) * I_solar / 4
I_atm_up = I_visible
I_atm_down = I_visible
I_ground_up = I_atm_up + I_atm_down - I_visible
T_ground = (I_ground_up / (epsilon * sigma))^0.25
```

T_ground = `r format_md(T_ground, digits = 3)` K. This is the same as the 
bare-rock temperature.

The effect of the dust in the atmosphere is to cancel out the greenhouse effect
and cool off the surface to the bare-rock temperature. 
The greenhouse effect works because the atmosphere is transparent to shortwave
light and opaque to longwave light. If the atmosphere becomes opaque to 
shortwave light, then the greenhouse effect doesn't work.

