---
title: Chapter 8 Homework Answers
hw_number: 5
pubdate: '2023-03-02'
date: '2023-03-01'
pdf_url: /files/homework_solutions/hw_05/homework_05_ch_8_answers.pdf
slug: homework_05_ch_8_answers
output:
  blogdown::html_page:
    md_extensions: +tex_math_single_backslash+compact_definition_lists
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 3.0
---
```{r setup, include=FALSE}
knitr::knit_hooks$set(inline = function(x) { knitr:::format_sci(x, 'md')})
knitr::opts_chunk$set(echo = FALSE, include = TRUE, cache = FALSE)

if (knitr::is_latex_output()) {
  message("Setting PDF plot hook.")
  new_plot_hook <- function(x, options)
    paste("\n", knitr::hook_plot_tex(x, options), "\n")

  knitr::knit_hooks$set(plot = new_plot_hook)
} else {
  message("Not using PDF output.")
}


options(dplyr.summarise.inform = FALSE)

# This section loads necessary R libraries and sources scripts that define
# useful functions format_md.
#
data_dir = "_data"
script_dir = "_scripts"

library(pacman)
p_load(tidyverse, jsonlite, httr, magrittr, scales, xml2, knitr)

theme_set(theme_bw(base_size = 10))

# Create a data directory if one does not exist.
if (!dir.exists(data_dir)) dir.create(data_dir)
if (!dir.exists(file.path(data_dir, "ch_08")))
    dir.create(file.path(data_dir, "ch_08"))


source(file.path(script_dir, "utils.R"), chdir = T)
source(file.path(script_dir, "format_md.R"), chdir = T)
source(file.path(script_dir, "geocarb.R"), chdir = T)
```
## Exercise 8.1: Weathering

> Weathering as a function of CO~2~. In steady state, the rate of weathering
> must balance the rate of CO~2~ degassing from the Earth, from volcanoes and
> deep sea vents. Run a simulation with a higher CO~2~ degassing rate at the
> transition time.
>
> a) Does an increase in CO~2~ degassing drive atmospheric CO~2~ up or down?
>    How long does this take?

### Answer to 8.1(a)

It takes about 1 million years for CO~2~ to come to equilibrium after the
volcanic degassing rate changes.

```{r answer-1a}
#| echo=FALSE, warning=FALSE, message=FALSE
ans_1a_file <- file.path(data_dir, "ch_08", "ex_1a.txt")
if (file.exists(ans_1a_file)) {
  ex_1a_data <- read_geocarb(ans_1a_file)
} else {
  ex_1a_data <- run_geocarb(filename = ans_1a_file, co2_spike = 0, 
                            degas_sim = 10)
}

ex_1a_data %>% filter(year <= 1E6) %>%
  ggplot(aes(x = year, y = co2_atmos)) +
  geom_line(size = 1, color = "darkblue") +
  scale_y_continuous(limits = c(0, 800), expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0), labels = label_comma(1)) +
  labs(x = "Year", y = expression(paste(CO[2], " (ppm)"))) +
  theme(plot.margin = margin(7, 25, 5, 5, unit = "pt"))
```

> b) How can you see that the model balances weathering against CO~2~
>    degassing?

### Answer to 8.1(b)

Look at the "Silicate Thermostat" graph: Watch how the weathering rate 
gradually rises until it meets the degassing rate.

```{r answer-1b}
#| echo=FALSE, warning=FALSE, message=FALSE
ex_1a_data %>% select(year, silicate_weathering, degassing_rate) %>%
  filter(year <= 1E6) %>%
  pivot_longer(-year, names_to = "var", values_to = "value") %>%
  ggplot(aes(x = year, y = value, color = var)) +
  geom_line(size = 1) +
  scale_color_manual(values = c(silicate_weathering = "darkblue",
                                degassing_rate = "darkred"),
                     labels = c(silicate_weathering = "Silicate Weathering",
                                degassing_rate = "Degassing Rate"),
                     name = NULL) +
  scale_x_continuous(expand = c(0,0),
                     labels = label_comma(1)) +
  scale_y_continuous(limits = c(0, 10.05), expand = c(0,0)) +
  labs(x = "Year", y = NULL) +
  theme(legend.position = c(0.99, 0.9),
        legend.justification = c(1, 1),
        legend.key.size = unit(2, "lines"),
        legend.background = element_rect(fill = "white", color = "black",
                                         linewidth = 0.2),
        plot.margin = margin(7, 25, 5, 5, unit = "pt"))
```

> c) Repeat this run with a range of degassing rates, and make a table of the
>    equilibrium CO~2~ concentration as a function of the CO~2~ degassing
>    rate.
>    The CO~2~ degassing rate is supposed to balance the CO~2~ consumption
>    rate by silicate weathering.

### Answer to 8.1(c)

```{r answer-1c}
#| echo=FALSE, warning=FALSE, message=FALSE
runs <- list()
for (degas in c(2.5, 5.0, 7.5, 10.0, 12.5, 15.0)) {
  file <- file.path(data_dir, "ch_08", 
                    str_c("ex_1c_", sprintf("%.1f", degas) %>% 
                            str_replace(fixed("."), "_"),
                          ".txt")
  )
  if (file.exists(file)) {
    df <- read_geocarb(file)
  } else {
    df <- run_geocarb(filename = file, co2_spike = 0, degas_sim = degas)
  }
  runs <- c(runs, list(df))
}

equil <- runs %>% map_df(~.x %>% filter(year == 1E6) %>% 
                           select(year, degas = degassing_rate,
                                  co2 = co2_atmos,
                                  weathering = silicate_weathering))
```
```{r answer-1c-table}
equil %>% select(degas, co2) %>% 
  knitr::kable(col.names = c(degas = "Degassing rate",
                                    co2 = "CO2"),
               digits = c(1,0))
```
```{r answer-1c-plot}
ggplot(equil, aes(x = degas, y = co2)) +
  geom_point(size = 3, color = "darkblue") + 
  geom_line(size = 1, color = "darkblue") +
  labs(x = "Degassing rate", y = expression(paste(CO[2], " (ppm)")))
```

> d) Make a plot of weathering as a function of atmospheric $p \COO$ using your
>    model runs.

### Answer to 8.1(d)

```{r answer-1d-plot}
ggplot(equil, aes(x = degas, y = weathering)) +
  geom_point(size = 3, color = "darkblue") + 
  geom_line(size = 1, color = "darkblue") +
  labs(x = "Degassing rate", y = "Silicate weathering rate")
```

## Exercise 8.2: Weathering as a Feedback

> Effect of solar intensity on steady state CO~2~ concentration. The rate of
> weathering is a function of CO~2~ and sunlight, a positive function of both
> variables. By this I mean that an increase in CO~2~ will drive an increase
> in weathering, as will an increase in sunlight. The sun used to be less
> intense than it is now. Turn back the clock 500 million years to when the sun
> was cooler than today. What do you get for the steady-state CO~2~, and how
> does this compare with what you get for today’s solar intensity? Explain why.

### Answer to 8.2

Using today's solar intensity, the CO~2~ concentration is 272.6 parts per 
million.

Using the solar intensity from 500 million years ago, we find a CO~2~
concentration of 2,372 parter per million, almost 10 times greater.

This is because, with less sunlight and the same CO~2~ concentration we
have today, the temperature would be much lower. The reduced temperature would 
slow down weathering and with less weathering the CO~2~ would build up in the 
atmosphere and increase the temperature until the weathering matches the 
degassing rate.
